DeepPicar: A Low-cost Deep Neural Network-based Autonomous Car

### 2018-04-20 paper/comparison.tex
Paper revisions

Figure~\ref{fig:sys_core} shows the average control loop completion
timing of the four system configurations we tested as a function of
the number of CPU cores used. (cf. Figure~\ref{fig:perf-vs-corecnt})
Both the Intel UP and Jetson TX2 exhibit better performance than
Raspberry Pi 3. When all four CPU cores are used, Intel UP is
1.33X faster than Pi 3, while TX2 (CPU) and TX2 (GPU) are 2.79X and
4.16X times faster, respectively, than Pi 3.
Thus, they all satisfy 33.3 ms WCET by a clear margin, and, in the
case of the TX2, 50 Hz or even 100 Hz real-time control is feasible
with the help of its GPU.
Another observation is that the DNN task's performance on TX2 (GPU)
does not change much as we increase the number of cores. This
isbecause most of the neural network computation is done by the GPU.

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/compare_model}
  \caption{Timing impact of co-scheduling multiple DNNs on different
    embedded multicore platforms. %% 1Nx1C: one DNN
    %% model using one core; 4Nx1C: four DNN models each using one core;
    %% 1Nx2C: one DNN model using two cores; 2Nx2C: two DNN models each
    %% using two cores.
  }
  \label{fig:sys_model}
\end{figure}

Figure~\ref{fig:sys_model} shows the results of the
multi-model co-scheduling experiment
(cf. Figure~\ref{fig:perf-vs-modelcnt}). Once again, they can comfortably
satisfy 30 Hz real-time performance for all of the co-scheduled DNN control
loops, and in the case of the TX2 (GPU), even 100 Hz real-time control
is feasible in all co-scheduling setups.
Given that the GPU must be shared among the co-scheduled DNN
models, the results suggest that the TX2's GPU has sufficient capacity to
accomodate multiple instances of the DNN models we tested.

\begin{figure}[h]
  \centering
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth]{figs/compare_benchmark_read}
    \caption{BwRead}
    \label{fig:sys_bench_read}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth]{figs/compare_benchmark}
    \caption{BwWrite}
    \label{fig:sys_bench_write}    
  \end{subfigure}
  \caption{Average processing time vs. the number of memory
    memory intensive co-runners.}
  \label{fig:sys_bench}    
\end{figure}

Figure~\ref{fig:sys_bench} shows the results of synthetic memory
intensive task co-scheduling experiment
(cf. Figure~\ref{fig:perf_vs_bandwidth}).
For read co-runners (BwRead), the performance of all platforms
gradually decreased as additional BwRead instances are introduced: up
to 1.6X for Pi 3, up to 1.3X for Intel UP, and up to 1.6X and 2.2X for
the TX2 (CPU) and TX2(GPU), respectively.
For write co-runners (BwWrite), however, we observe generally more
noticieable execution time increases. As we discussed
earlier in Section~\ref{sec:evaluation}, Pi 3 suffers up to 11.6X
execution time increase, while Intel UP and Jetson TX2 suffer less
dramatic but still significant execution time increases.

@@

Another interesting observation is that the TX2 (GPU) also suffers
considerable execution time increase (2.3X) despite the fact that the
co-scheduled synthetic tasks do not utilize the GPU (i.e., gthe DNN
model has dedicated access to the GPU.) This is, however, a 
known characteristic of integrated CPU-GPU architecture based
platforms in which both CPU and GPU share the same memory
subsystem~\cite{Ali2017} and therefore can suffer bandwidth contention
as we observe in this case.

@@

In summary, we find that todays embedded computing platforms, even as
inexpensive as a Raspberry Pi 3, are powerful enough to support
DNN based real-time control applications. Furthermore, availability of
CPU cores and GPU on these platforms allow consolidating mutiple DNN
workloads. However, shared resource contention among these diverse
computing resources remains an important issue that must be understood
and controlled, especially for safety-critical applications.

>>>>

Figure~\ref{fig:sys_core} shows the average control loop completion
timing of the four system configurations we tested as a function of
the number of CPU cores used. (cf. Figure~\ref{fig:perf-vs-corecnt})
Both the Intel UP and Jetson TX2 exhibit better performance than
Raspberry Pi 3. When all four CPU cores are used, the Intel UP is
1.33X faster than Pi 3, while the TX2 (CPU) and TX2 (GPU) are 2.79X and
4.16X times faster that the Pi 3, respectively.
Thus, they all satisfy 33.3 ms WCET by a clear margin, and, in the
case of the TX2, 50 Hz or even 100 Hz real-time control is feasible
with the help of its GPU.
Another observation is that the CNN task's performance on TX2 (GPU)
does not change much as we increase the number of cores. This
is because most of the neural network computation is done by the GPU.

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/compare_model}
  \caption{Timing impact of co-scheduling multiple CNNs on different
    embedded multicore platforms. %% 1Nx1C: one DNN
    %% model using one core; 4Nx1C: four DNN models each using one core;
    %% 1Nx2C: one DNN model using two cores; 2Nx2C: two DNN models each
    %% using two cores.
  }
  \label{fig:sys_model}
\end{figure}

Figure~\ref{fig:sys_model} shows the results of the
multi-model co-scheduling experiment
(cf. Figure~\ref{fig:perf-vs-modelcnt}). Once again, they can comfortably
satisfy 30 Hz real-time performance for all of the co-scheduled CNN control
loops, and in the case of the TX2 (GPU), even 100 Hz real-time control
is feasible in all co-scheduling setups.
Given that the GPU must be shared among the co-scheduled CNN
models, the results suggest that the TX2's GPU has sufficient capacity to
accomodate multiple instances of the CNN models we tested.

\begin{figure}[h]
  \centering
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth]{figs/compare_benchmark_read}
    \caption{BwRead}
    \label{fig:sys_bench_read}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.45\textwidth}
    \includegraphics[width=\textwidth]{figs/compare_benchmark}
    \caption{BwWrite}
    \label{fig:sys_bench_write}    
  \end{subfigure}
  \caption{Average processing time vs. the number of memory
    intensive co-runners.}
  \label{fig:sys_bench}    
\end{figure}

Figure~\ref{fig:sys_bench} shows the results of the synthetic memory
intensive task co-scheduling experiments
(cf. Figure~\ref{fig:perf_vs_bandwidth}).
For read co-runners (BwRead), the performance of all platforms
gradually decreased as additional BwRead instances were introduced: up
to 1.6X for the Pi 3, up to 1.3X for the Intel UP, and up to 1.6X and 2.2X for
the TX2 (CPU) and TX2(GPU), respectively.
For write co-runners (BwWrite), however, we observe generally more
noticieable execution time increases. As we discussed
earlier in Section~\ref{sec:evaluation}, the Pi 3 suffers up to 11.6X
execution time increase, while the Intel UP and Jetson TX2 suffer less
dramatic but still significant execution time increases.

@@

Another interesting observation is that the TX2 (GPU) also suffers
considerable execution time increase (2.3X) despite the fact that the
co-scheduled synthetic tasks do not utilize the GPU (i.e., the CNN
model has dedicated access to the GPU.) This is, however, a 
known characteristic of integrated CPU-GPU architecture based
platforms in which both the CPU and GPU share the same memory
subsystem~\cite{Ali2017} and therefore can suffer bandwidth contention
as we observe in this case.

@@

In summary, we find that todays embedded computing platforms, even as
inexpensive as a Raspberry Pi 3, are powerful enough to support
CNN based real-time control applications. Furthermore, availability of
CPU cores and a GPU on these platforms allows consolidating multiple CNN
workloads. However, shared resource contention among these diverse
computing resources remains an important issue that must be understood
and controlled, especially for safety-critical applications.

### 2018-04-20 paper/conclusion.tex
Paper revisions

Despite the complexity of the neural network, DeepPicar uses a
low-cost embedded quad-core computer, the Raspberry Pi 3,  as its sole
computing resource.
We systematically analyzed the platform's real-time capability in
supporting the DNN-based real-time control task.
We also evaluated other, more powerful, embedded computing
platforms to better understand achievable real-time performance of
DeepPicar's DNN based control system and the impact of
computing hardware architectures.
We find all tested embedded platforms, including the Pi 3, are capable
of supporting deep-learning based real-time control, from 20 Hz up to
100 Hz, depending on the platform and its system
configuration. Futhermore, all platforms were capable of consolidating
multiple DNN models and/or tasks.

However, when consolidating multiple models/tasks, shared resource
contention remains an important issue that must be considered to
ensure desired real-time performance on these shared memory based
embedded computing platforms. 
Toward this end, we evaluated the impact of shared resource contention
to the DNN workload in diverse consolidated workload setups, and
evaluated the effectivness of state-of-the-art shared resource
isolation mechanisms in protecting real-time performance of the DNN
workload.

%% As future work, we plan to improve the prediction accuracy of the
%% system by feeding more data in training.
As future work, we plan to investigate ways to reduce computational
and memory overhead of DNN inferencing and to evaluate the
effectiveness of FPGA and other specialized accelerators.

>>>>

Despite the complexity of the neural network, DeepPicar uses a
low-cost embedded quad-core computer, the Raspberry Pi 3,  as its sole
computing resource.
We systematically analyzed the platform's real-time capability in
supporting the CNN-based real-time control task.
We also evaluated other, more powerful, embedded computing
platforms to better understand achievable real-time performance of
DeepPicar's CNN based control system and the impact of
computing hardware architectures.
We find all tested embedded platforms, including the Pi 3, are capable
of supporting deep-learning based real-time control, from 20 Hz up to
100 Hz, depending on the platform and its system
configuration. Futhermore, all platforms were capable of consolidating
multiple CNN models and/or tasks.

However, when consolidating multiple models/tasks, shared resource
contention remains an important issue that must be considered to
ensure desired real-time performance on these shared memory based
embedded computing platforms. 
Toward this end, we evaluated the impact of shared resource contention
to the CNN workload in diverse consolidated workload setups, and
evaluated the effectivness of state-of-the-art shared resource
isolation mechanisms in protecting real-time performance of the CNN
workload.

%% As future work, we plan to improve the prediction accuracy of the
%% system by feeding more data in training.
As future work, we plan to investigate ways to reduce computational
and memory overhead of CNN inferencing and to evaluate the
effectiveness of FPGA and other specialized accelerators.

### 2018-04-20 paper/evaluation-cachepart.tex
Paper revisions

Cache partitioning is a well-known technique to improve isolation in
multicore by giving a dedicated cache space to each individaul 
task or core. In this experiment, we use PALLOC~\cite{yun2014rtas}, a
page-coloring based kernel-level memory allocator for Linux.
Page coloring is a OS technique that controls the physical addresses
of memory pages. By allocating pages over non-overlapping sets of the
cache, the OS can effectively partition the cache.
Using PALLOC, we investigate the effect of cache partitioning in
protecting DeepPicar's DNN based controller.

@@

Figure~\ref{fig:cache-mapping} shows the physical address
mapping of Raspberry Pi 3's BCM2837 processor, which has 32K private
L1 I\&D (4way) caches and a share 512KB L2 (16 way) cache. In order not
to partition the private L1 caches, we use the bit 13 and 14 for
coloring, which results in 4 usuable page colors.

@@

In the first experiment, we investigate the cache space sensitivity of
DeepPicar's DNN-based control loop. Using PALLOC, we create 4
different cgroups which are configured to use 4, 3,
2, and 1 colors (100\%, 75\%, 50\% and 25\% of the L2 cache
space, respectively). We then execute the DNN control loop (inference)
on one core using a different cgroup cache partition, one at a time,
and measure the average processing time.

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/palloc_multicore}
  \caption{Cache space sensitivity of the DNN controller.}
  \label{fig:palloc_multicore}
\end{figure}

Figure~\ref{fig:palloc_multicore} shows the result. As can be seen,
the DNN inference timing is hardly changed at all regardless of the
size of the allocated L2 cache space. In other words, we find that
the DNN workload is largely insensitive to L2 cache space.

The next experiment further validates this finding. In this
experiment, we repeat the experiment in
Section~\ref{sec:eval-memhog}---i.e., co-scheduling the DNN model and
three Bandwidth (BwRead or BwWrite) instances---but this time we
ensure that each task is given equal amounts of L2 cache space by
assigning one color to each task's cache partition.

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/palloc_bandwidth_exectime}
  \caption{Average processing time vs. the number of memory
memory intensive co-runners; Each core (task) is given a equal-sized
dedicated cache partition.}
  \label{fig:palloc_bandwidth_exectime}
\end{figure}

Figure \ref{fig:palloc_bandwidth_exectime} shows the
result. For BwRead, when cache partitioning is enabled, DNN
performance becomes slightly better, with a improvement of  
\textasciitilde6.5\%. For BwWrite, cache partitioning actually makes
the DNN performance slightly worse that that of w/o partitioning.

In summary, we find that the DNN inferencing workload is not sensitive
to cache space and that cache partitioning is not effective in
providing timing isolation to our DNN workload.

>>>>

Cache partitioning is a well-known technique to improve isolation in
multicore systems by giving a dedicated cache space to each individaul 
task or core. In this experiment, we use PALLOC~\cite{yun2014rtas}, a
page-coloring based kernel-level memory allocator for Linux.
Page coloring is a OS technique that controls the physical addresses
of memory pages. By allocating pages over non-overlapping cache sets,
the OS can effectively partition the cache.
Using PALLOC, we investigate the effect of cache partitioning on
protecting DeepPicar's CNN based controller.

@@

Figure~\ref{fig:cache-mapping} shows the physical address
mapping of the Raspberry Pi 3's BCM2837 processor, which has 32K private
L1 I\&D (4way) caches and a shared 512KB L2 (16 way) cache. In order
to avoid partitioning the private L1 caches, we use bits 13 and 14 for
coloring, which results in 4 usuable page colors.

@@

In the first experiment, we investigate the cache space sensitivity of
the DeepPicar's CNN-based control loop. Using PALLOC, we create 4
different cgroups which are configured to use 4, 3,
2, and 1 colors (100\%, 75\%, 50\% and 25\% of the L2 cache
space, respectively). We then execute the CNN control loop (inference)
on one core using a different cgroup cache partition, one at a time,
and measure the average processing time.

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/palloc_multicore}
  \caption{Cache space sensitivity of the CNN controller.}
  \label{fig:palloc_multicore}
\end{figure}

Figure~\ref{fig:palloc_multicore} shows the results. As can be seen,
the CNN inference timing hardly changes at all regardless of the
size of the allocated L2 cache space. In other words, we find that
the CNN workload is largely insensitive to L2 cache space.

The next experiment further validates this finding. In this
experiment, we repeat the experiment in
Section~\ref{sec:eval-memhog}---i.e., co-scheduling the CNN model and
three Bandwidth (BwRead or BwWrite) instances---but this time we
ensure that each task is given equal amounts of L2 cache space by
assigning one color to each task's cache partition.

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/palloc_bandwidth_exectime}
  \caption{Average processing time vs. the number of memory
intensive co-runners; Each core (task) is given a equal-sized
dedicated cache partition.}
  \label{fig:palloc_bandwidth_exectime}
\end{figure}

Figure \ref{fig:palloc_bandwidth_exectime} shows the
results. Compared to Figure~\ref{fig:perf_vs_bandwidth} where no cache
partioning is applied, assigning a dedicated L2 cache parititon to
each core does not provide significant isolation benefits. For BwRead
co-runners, cache partitioning slightly improves isolation, but for
BwWrite co-runners, cache partitioning causes worse worst-case
slowdown.

In summary, we find that the CNN inferencing workload is not sensitive
to cache space and that cache partitioning is not effective in
providing timing isolation for our CNN workload.

### 2018-04-20 paper/evaluation-drampart.tex
Paper revisions

In this subsection, we examine the DNN workload's memory bandwidth
sensitivity and the effect of memory bandwidth throttling in providing
isolation. For the experiments, we use MemGuard\cite{Yun2013}, which 
can limit the amount of memory bandwidth each core receives. MemGuard
operates periodically, at an 1ms interval, and uses hardware
performance counters to throttle cores if they exceed their given
bandwidth budgets within each regulation period (i.e., 1ms) by
scheduling high-priority idle kernel threads, until the next period begins.

In the first experiment, we measure the performance of the DNN model
on a single core, first w/o using MemGuard and then w/
using MemGuard while varying the core's bandwidth throttling parameter
from 500 MB/s down to 100 MB/s.

@@

Figure \ref{fig:memguard_multicore} shows the result. When core
executing the DNN model is throttled at 400 MB/s or more, performance
of the model is largely the same as the non-throttled case. However, as
we decrease the assigned memory bandwidth below 300 MB/s, we start to
observe noticeable decreases in the model's performance. In other
words, the DNN model is sensitive to memory bandwidth and that it
requires 400 MB/s or more bandwidth should be given to ensure ideal
performance.

In the next experiment, we again repeat the experiment in
Section~\ref{sec:eval-memhog}---i.e., co-scheduling memory intensive
synthetic tasks---but this time we
throttle the cores of the co-runners using MemGuard and vary their
memory bandwidth budgets to see their impact on the DNN model's 
performance.

@@

Figures \ref{fig:memguard_bandwidth} shows the result.
Note that, in case of BwWrite co-runners, we have assigned half the
bandwidth budget because MemGuard currently only account L2 refills
but not write-backs, which effectively allow twice the allocated
memory bandwidth budget. Regardless, the result clearly shows that
limiting co-runners's memory bandwidth is effective in protecting the
DNN model's performance.

In summary, we find that the DNN inferencing workload is sensitive to
memory bandwidth and that memory bandwidth throttling is effective in
improving performance isolation of the DNN workload.

>>>>

In this subsection, we examine the CNN workload's memory bandwidth
sensitivity and the effect of memory bandwidth throttling in providing
isolation. For the experiments, we use MemGuard\cite{Yun2013}, which 
can limit the amount of memory bandwidth each core receives. MemGuard
operates periodically, at a 1 ms interval, and uses hardware
performance counters to throttle cores if they exceed their given
bandwidth budgets within each regulation period (i.e., 1 ms) by
scheduling high-priority idle kernel threads, until the next period begins.

In the first experiment, we measure the performance of the CNN model
on a single core, first w/o using MemGuard and then w/
using MemGuard while varying the core's bandwidth throttling parameter
from 500 MB/s down to 100 MB/s.

@@

Figure \ref{fig:memguard_multicore} shows the results. When the core
executing the CNN model is throttled at 400 MB/s or more, the performance
of the model is largely the same as the non-throttled case. However, as
we decrease the assigned memory bandwidth below 300 MB/s, we start to
observe noticeable decreases in the model's performance. In other
words, the CNN model is sensitive to memory bandwidth and it
requires 400 MB/s or more bandwidth to ensure ideal performance.

In the next experiment, we repeat the experiment in
Section~\ref{sec:eval-memhog}---i.e., co-scheduling memory intensive
synthetic tasks---but this time we
throttle the cores of the co-runners using MemGuard and vary their
memory bandwidth budgets to see their impact on the CNN model's 
performance.

@@

Figures \ref{fig:memguard_bandwidth} shows the results.
Note that, in case of BwWrite co-runners, we have assigned half the
bandwidth budget because MemGuard currently only accounts for L2 refills
but not write-backs, which effectively allows twice the allocated
memory bandwidth budget. Regardless, the results clearly show that
limiting the co-runners's memory bandwidth is effective in protecting the
CNN model's performance.

In summary, we find that the CNN inferencing workload is sensitive to
memory bandwidth and that memory bandwidth throttling is effective in
improving the performance isolation of the CNN workload.

### 2018-04-20 paper/evaluation.tex
Paper revisions

\subsection{Setup}
The Raspberry Pi 3 Model B platform used in DeepPicar equips a Broadcom
BCM2837 SoC, which has a quad-core ARM Cortex-A53 cluster,
running at up to 1.2GHz. Each core has private 32K I/D caches
and all cores share a 512KB L2 cache.
The chip also includes Broadcom's Videocore IV GPU, although we did
not use the GPU in our evaluation, due to the lack of sofware support
\footnote{TensorFlow currently only supports NVIDIA's GPUs.}.
For software, we use Ubuntu MATE 16.04, TensorFlow 1.1.
We disabled DVFS (dynamic voltage frequency scaling) and
configured the clock speed of each core statically at the maximum 1.2GHz.
We use the SCHED\_FIFO real-time scheduler to schedule the DNN control
task while using the CFS when executing memory intensive co-runners.

\subsection{Inference Timing for Real-Time Control}
For real-time control of a car (or any robot), the control loop
frequency must be sufficiently high so that the car can quickly
react to the changing environment and its internal states. In general,
control performance improves when the frequency is higher, though
computation time and the type of particular physical system are
factors in determining a proper control loop frequency. While a standard
control system may be comprised of multiple control loops with
differing control frequencies---e.g., an inner control loop for lower-level
PD control, an outer loop for motion planning, etc.---DeepPicar's
control loop is a single layer, as shown earlier in
Figure~\ref{fig:controlloop}, since a single deep neural network
replaces the traditional multi-layer control pipline. (Refer to
Figure~\ref{fig:end-to-end-control} on the differences between the
standard robotics control vs. end-to-end deep learning approach).
This means that the DNN inference operation must be completed
within the inner-most control loop update frequency.

To understand achievable control-loop update frequencies, we
experimentally measured the execution times of DeepPicar's DNN
inference operations.

@@

\begin{table}[h]
  \centering
  \begin{tabular} {| c | r | r | r | r |}
    \hline
    \textbf{Operation} & \textbf{Mean} & \textbf{Max} &   \textbf{99pct.} & \textbf{Stdev.} \\ \hline
    Image capture        & 1.61  &  1.81 &  1.75  & 0.05 \\ \hline
    Image pre-processing & 2.77  &  2.90 &  2.87  & 0.04 \\ \hline
    DNN inferencing      & 18.49 & 19.30 & 18.99  & 0.20 \\ \hline
    Total Time           & 22.87 & 23.74 & 23.38  & 0.20 \\ \hline
  \end{tabular}
  \caption{Control loop timing breakdown.}
  \label{tbl:control-loop-breakdown}
\end{table}

Figure~\ref{fig:control-loop-timing} shows the measured control loop 
processing times of the DeepPicar over 1000 image frames (one per each
control loop). We omit the first frame's processing time for cache
warmup. Table~\ref{tbl:control-loop-breakdown} shows the time
breakdown of each control loop. Note that all four CPU cores of the
Raspberry Pi 3 were used by the TensorFlow library when performing the
DNN inference operations.

@@

Lastly, we find that the control loop execution timing is highly
predictable and shows very little variance over different input image
frames. This is because the amount of computation needed to perform
a DNN inferencing operation is fixed at the DNN architecture design
time and does not change at runtime over different inputs (i.e.,
different image frames). This predictable timing behavior is a highly
desirable property for real-time systems, making DNN inferencing an
attractive real-time workload.

@@

As noted in~\cite{NVIDIA2015}, DNN inferencing is inherently more
difficult to parallelize than training because the easiest
parallelization option, batching (i.e., processing multiple images in
parallel), is not available or is limited. Specifically, in DeepPicar,
only one image frame, obtained from the camera, can be processed at a
time. Thus, more fine-grained algorithmic parallelization is needed to
improve inference performance~\cite{NVIDIA2015}, which generally does
not scale well.

@@

\subsection{Effect of Co-scheduling Multiple DNN Models}

In this experiment, we launch multiple instances of DeepPicar's DNN
model at the same time and measure its impact on their inference
timings. In other words, we are interested in how shared resource
contention affects inference timing. These instances are not identical
copies of the same model, but are instead copies of different models
(but of the same network architecture). 
This is done to ensure that no L2 cache memory is shared between the 
models as they are running.

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/perf_vs_modelcnt}
  \caption{Timing impact of co-scheduling multiple DNNs. 1Nx1C: one DNN
    model using one core; 4Nx1C: four DNN models each using one core;
    1Nx2C: one DNN model using two cores; 2Nx2C: two DNN models each
    using two cores.} 
  \label{fig:perf-vs-modelcnt}
\end{figure}

Figure~\ref{fig:perf-vs-modelcnt} shows the results. In the figure, the
X-axis shows the system configuration: \#of DNN models x \#of CPU
cores/DNN. For example, `4Nx1C' means running four DNN models each of
which is assigned to run on one CPU core, whereas `2Nx2C' means running
two DNN models, each of which is assigned to run on two CPU
cores. The Y-axis shows the average inference timing.
The two bars on the left show the impact of co-scheduling four DNN
models. Compared to executing a single DNN model on one CPU core
(1Nx1C), when four DNN models are co-scheduled (4Nx1C), each model
suffers an average inference time increase of approximately 11 ms,
or 24\%. On the other hand, when two DNN models, each using two CPU
cores, are co-scheduled (2Nx2C), the average inference timing is increased by
about 4 ms, or 13\%, compared to the baseline of running one model
using two CPU cores (1Nx2C).

@@

In this experiment, we investigate \emph{worst-case} impact of shared 
resource contention to DeepPicar's DNN inference timing using  
a synthetic memory benchmark. Specifically, we use the \emph{Bandwidth}
benchmark from the IsolBench suite~\cite{Valsan2016}, which
sequentially reads or writes a big array; we henceforth refer BwRead 
as Bandwidth with read accesses and BwWrite as the one with write
accesses. The experiment setup is as follows. We run a single DNN
model on one core, and co-schedule an increasing number of the
Bandwidth benchmark instances on the other cores. We repeat the
experiement first with BwRead and next with BwWrite.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/perf_vs_bandwidth}
  \caption{Average processing time vs. the number of memory
    memory intensive co-runners introduced.}
  \label{fig:perf_vs_bandwidth}
\end{figure}

@@

Figure~\ref{fig:perf_vs_bandwidth} shows the result. Note
that BwWrite co-runners cause significantly high execution time
incrases---up to 11.6X---on DNN inferencing while BwRead co-runners
cause relatively much smaller time increases. While execution time
increases are expected, the degree of which in the worst-case is quite
surprising---\emph{if the DNN controller was driving an actual car, it
  would result in a crash of the car!}

Given the importance of predictable timing on real-time control, such
as our DNN based control task, we wanted to understand (1) the main
source of the timing and (2) to evaluate existing isolation methods to
avoid this kind of timing interference. Specifically, shared cache
space and DRAM bandwidth are two most well-known sources of contention
in multicore. Thus, in the following, we investigate if and how
they influence the observed timing interference and how effective are
the existing mitigation methods.

>>>>

\subsection{Setup}
The Raspberry Pi 3 Model B platform used in DeepPicar equips a Broadcom
BCM2837 SoC, which has a quad-core ARM Cortex-A53 cluster,
running at up to 1.2GHz. Each core has private 32K I/D caches
and all cores share a 512KB L2 cache.
The chip also includes Broadcom's Videocore IV GPU, although we did
not use the GPU in our evaluation, due to the lack of sofware support
\footnote{TensorFlow currently only supports NVIDIA's GPUs.}.
For software, we use Ubuntu MATE 16.04 and TensorFlow 1.1.
We disabled DVFS (dynamic voltage frequency scaling) and
configured the clock speed of each core statically at the maximum 1.2GHz.
We use the SCHED\_FIFO real-time scheduler to schedule the CNN control
task while using the CFS when executing memory intensive co-runners.

\subsection{Inference Timing for Real-Time Control}
For real-time control of a car (or any robot), the control loop
frequency must be sufficiently high so that the car can quickly
react to the changing environment and its internal states. In general,
control performance improves when the frequency is higher, though
computation time and the type of particular physical system are
factors in determining a proper control loop frequency. While a standard
control system may be comprised of multiple control loops with
differing control frequencies---e.g., an inner control loop for lower-level
PD control, an outer loop for motion planning, etc.---DeepPicar's
control loop is a single layer, as shown earlier in
Figure~\ref{fig:controlloop}, since a single deep neural network
replaces the traditional multi-layer control pipline. (Refer to
Figure~\ref{fig:end-to-end-control} on the differences between the
standard robotics control vs. end-to-end deep learning approach).
This means that the CNN inference operation must be completed
within the inner-most control loop update frequency.

To understand achievable control-loop update frequencies, we
experimentally measured the execution times of DeepPicar's CNN
inference operations.

@@

\begin{table}[h]
  \centering
  \begin{tabular} {| c | r | r | r | r |}
    \hline
    \textbf{Operation} & \textbf{Mean} & \textbf{Max} &   \textbf{99pct.} & \textbf{Stdev.} \\ \hline
    Image capture        & 1.61  &  1.81 &  1.75  & 0.05 \\ \hline
    Image pre-processing & 2.77  &  2.90 &  2.87  & 0.04 \\ \hline
    CNN inferencing      & 18.49 & 19.30 & 18.99  & 0.20 \\ \hline
    Total Time           & 22.87 & 23.74 & 23.38  & 0.20 \\ \hline
  \end{tabular}
  \caption{Control loop timing breakdown.}
  \label{tbl:control-loop-breakdown}
\end{table}

Figure~\ref{fig:control-loop-timing} shows the measured control loop 
processing times of the DeepPicar over 1000 image frames (one per each
control loop). We omit the first frame's processing time for cache
warmup. Table~\ref{tbl:control-loop-breakdown} shows the time
breakdown of each control loop. Note that all four CPU cores of the
Raspberry Pi 3 were used by the TensorFlow library when performing the
CNN inference operations.

@@

Lastly, we find that the control loop execution timing is highly
predictable and shows very little variance over different input image
frames. This is because the amount of computation needed to perform
a CNN inferencing operation is fixed at the CNN architecture design
time and does not change at runtime over different inputs (i.e.,
different image frames). This predictable timing behavior is a highly
desirable property for real-time systems, making CNN inferencing an
attractive real-time workload.

@@

As noted in~\cite{NVIDIA2015}, CNN inferencing is inherently more
difficult to parallelize than training because the easiest
parallelization option, batching (i.e., processing multiple images in
parallel), is not available or is limited. Specifically, in DeepPicar,
only one image frame, obtained from the camera, can be processed at a
time. Thus, more fine-grained algorithmic parallelization is needed to
improve inference performance~\cite{NVIDIA2015}, which generally does
not scale well.

@@

\subsection{Effect of Co-scheduling Multiple CNN Models}

In this experiment, we launch multiple instances of DeepPicar's CNN
model at the same time and measure its impact on their inference
timings. In other words, we are interested in how shared resource
contention affects inference timing. These instances are not identical
copies of the same model, but are instead copies of different models
(but of the same network architecture). 
This is done to ensure that no L2 cache memory is shared between the 
models as they are running.

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/perf_vs_modelcnt}
  \caption{Timing impact of co-scheduling multiple CNNs. 1Nx1C: one CNN
    model using one core; 4Nx1C: four CNN models each using one core;
    1Nx2C: one CNN model using two cores; 2Nx2C: two CNN models each
    using two cores.} 
  \label{fig:perf-vs-modelcnt}
\end{figure}

Figure~\ref{fig:perf-vs-modelcnt} shows the results. In the figure, the
X-axis shows the system configuration: \#of CNN models x \#of CPU
cores/CNN. For example, `4Nx1C' means running four CNN models each of
which is assigned to run on one CPU core, whereas `2Nx2C' means running
two CNN models, each of which is assigned to run on two CPU
cores. The Y-axis shows the average inference timing.
The two bars on the left show the impact of co-scheduling four CNN
models. Compared to executing a single CNN model on one CPU core
(1Nx1C), when four CNN models are co-scheduled (4Nx1C), each model
suffers an average inference time increase of approximately 11 ms,
or 24\%. On the other hand, when two CNN models, each using two CPU
cores, are co-scheduled (2Nx2C), the average inference timing is increased by
about 4 ms, or 13\%, compared to the baseline of running one model
using two CPU cores (1Nx2C).

@@

In this experiment, we investigate the \emph{worst-case} impact of shared 
resource contention on DeepPicar's CNN inference timing using  
a synthetic memory benchmark. Specifically, we use the \emph{Bandwidth}
benchmark from the IsolBench suite~\cite{Valsan2016}, which
sequentially reads or writes a big array; we henceforth refer to BwRead 
as Bandwidth with read accesses and BwWrite as the one with write
accesses. The experiment setup is as follows: We run a single CNN
model on one core, and co-schedule an increasing number of the
Bandwidth benchmark instances on the other cores. We repeat the
experiement first with BwRead and next with BwWrite.

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/perf_vs_bandwidth}
  \caption{Average processing time vs. the number of memory
    intensive co-runners introduced.}
  \label{fig:perf_vs_bandwidth}
\end{figure}

@@

Figure~\ref{fig:perf_vs_bandwidth} shows the results. Note
that BwWrite co-runners cause significantly higher execution time
increases---up to 11.6X---on CNN inferencing while BwRead co-runners
cause relatively much smaller time increases. While execution time
increases are expected, the degree that is seen in the worst-case 
is quite surprising---\emph{if the CNN controller was driving an actual 
car, it would result in a car crash!}

Given the importance of predictable timing for real-time control, such
as our CNN based control task, we wanted to: (1) understand the main
source of the timing and (2) evaluate existing isolation methods to
avoid this kind of timing interference. Specifically, shared cache
space and DRAM bandwidth are the two most well-known sources of contention
in multicore systems. Thus, in the following sections, we investigate whether
and to what extent they influence the observed timing interference and 
the effectiveness of existing mitigation methods.
